---
title: "miRcore R camp TCGA Group Task"
author: "[Group Name Here]"
date: "August 8, 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```

# miRcore summer camp 2025
© miRcore


## Disease Information

What is the significance ((background and problems) of the disease you are analyzing?



## Reading in Data

Read in the cancer miRNA expression text file that you downloaded from the Google Drive

```{r}
LIHC = read.table("LIHC.txt", header = T, row.names = 1, stringsAsFactors = F)

```


A. What are the dimensions of your text file? 915 x 98

```{r}
dim(LIHC)
```

B. How many Tumor and Control samples are in your dataset? 49 and 49

```{r}
grep("Control", colnames(LIHC))
colnames(LIHC)

```


## Plotting

A1. Make box plots for first 10 Tumor samples and first 10 Control samples using all data.

```{r}

boxplot(LIHC[,c(1:10,50:60)])
```

A2. Make box plots for first 10 Tumor samples and first 10 Control samples using log2(data+1).

```{r}
LIHCM = as.matrix(LIHC)
log.LIHCM = log2(LIHCM+1)
boxplot(log.LIHCM[,c(1:10,50:60)])
```


B. Make a scatterplot for hsa-miR-200a-3p and hsa-miR-429 using all
samples in your data file. Then compute the Pearson correlation.

```{r}
LIHCM = as.matrix(LIHC)
plot(LIHCM["hsa-miR-200a-3p",],LIHCM["hsa-miR-429",])
cor(LIHCM["hsa-miR-200a-3p",],LIHCM["hsa-miR-429",])

```


## T Tests

A. For each miRNA: calculate mean, standard deviation, and the 2-tail student t-test p-value of Tumor and Control groups. Save all results in a table.

```{r}
ttable = data.frame()
control = grep("Control",colnames(LIHCM))
tumor = grep("Tumor",colnames(LIHCM))
for(j in 1:nrow(LIHCM)){
  
  meanControl =  mean(LIHCM[j,control])
  meanTumor = mean(LIHCM[j,tumor])
  sdControl = sd(LIHCM[j,control])
  sdTumor = sd(LIHCM[j,tumor])
  pval = t.test(LIHCM[j,tumor],LIHCM[j,control])$p.value
  temp = c(meanControl,meanTumor,sdControl,sdTumor,pval)
  ttable = rbind(ttable,temp)
}
colnames(ttable) = c("Control Mean", "Tumor Mean", "Control SD", "Tumor SD", "P-val")
row.names(ttable) = row.names(LIHC)




```


B.  Find 50 most significantly different miRNA names between tumor and control samples and save their names with means, standard deviations for tumor and control samples and p-values.
```{r}
sigvals = ttable[,"P-val"]
sigvec = order(sigvals)
sigvec = sigvec[1:50]
sigtable = ttable[sigvec,]
sigMIRNAS = rownames(sigtable)

sigtable
sigMIRNAS
```
```{r}
temptable = ttable[!is.nan(ttable[,"P-val"]),]
  for(l in 2:nrow(temptable)){
     for(z in (l-1):1){
       
        if(temptable[z,"P-val"] > temptable[z+1,"P-val"]){
          temp <- temptable[z,]
          temptable[z,] <- temptable[z+1,]
          temptable[z+1,] <- temp
      }
      
    }
  }
head(temptable)


```

```{r}

pvec = vector()
for(k in 1:nrow(ttable)){
  pvec = c(pvec, ttable[k,"P-val"])
  currentval = pvec[k]
  for(p in 1:length(pvec)){
    if(!is.na(pvec[p]) & !is.na(currentval) & pvec[p] > currentval){
      for(m in p:length(pvec)){
        temp = pvec[m]
        pvec[m] = currentval
        currentval = temp
      }
      
    }
  }
}
pvec[1:50]

```


C. Make a heatmap of 50 most signficantly differently expressed miRNAs (row clustering only). Use the colors darkblue,
white and red2. [USA!USA!USA!].

```{r}
install.packages("pheatmap")
library(pheatmap)

sigtable_LIHC <- log2(LIHC[sigMIRNAS,] +1)
sigtable_LIHC
pheatmap(sigtable_LIHC, scale = "row",clustering_distance_rows = "correlation", cellwidth = 15, cellheight = 10, color = colorRampPalette(c("darkblue", "white","red"))(256), border_color = NA, main = "Heatmap of top 50 sig Expressed",width = 100, height = 100)

```

## Correlations

A. Find correlations between the most significant miRNA with 29 other miRNAs for 1) all samples.


B. Among A. can you find some patterns between miRNAs? What are your findings? (example: to find the most correlated miRNA with the most significant miRNA)


## Random Forest (Bonus)

A. Create a random forest using the steps from the random forest .html file to classify your data. What is your accuracy?
```{r}


LIHC_data = read.table("LIHC.txt", header=T, row.names=1, stringsAsFactors=F)
head(LIHC_data)
dim(LIHC_data)

sig_miRs <- vector()

num_samples <- length(LIHC_data) # columns (172)
num_miRNAs <- nrow(LIHC_data) # rows (941)

tumor_samples <- 1:49 # 1:(num_samples/2) # 1:86
control_samples <- 50:98 # ((num_samples/2)+1):(num_samples) # 87:172

for(i in 1:num_miRNAs){
  tumor_vector <- LIHC_data[i,tumor_samples]
  control_vector <- LIHC_data[i,control_samples]
  
  p_value <- t.test(tumor_vector,control_vector)$p.value
  
  if(!is.na(p_value) && p_value < 0.001){
    sig_miRs <- c(sig_miRs,i)
  }
}

sig_miRs[1:10] # 4  6  8  9 10 11 16 17 18 22
length(sig_miRs)


LIHC_data_cleaned = LIHC_data[rowSums(LIHC_data > 0) > 86, ]
dim(LIHC_data_cleaned)


pval=vector()
for (i in 1:nrow(LIHC_data_cleaned)) {
  p_value = t.test(LIHC_data_cleaned[i,1:49],LIHC_data_cleaned[i,50:98])$p.value
  pval=c(pval,p_value)
}

top50 = order(pval)[1:50]
LIHC_top50 = LIHC_data_cleaned[top50,]

LIHC_data_sig = LIHC_data[sig_miRs,]

dim(LIHC_data_sig)

tLIHC_data_sig = t(LIHC_data_sig)

Condition = rownames(tLIHC_data_sig)
Condition[1:10]

for(i in 1:length(Condition)){
  if (grepl("Tumor", Condition[i], fixed = TRUE) == TRUE) {
     Condition[i] = "Tumor"
  } else{
     Condition[i] = "Control"
  }
}

Condition = as.factor(Condition)
tLIHC_data_sig = cbind.data.frame(Condition, tLIHC_data_sig)
tLIHC_data_sig[1:10,1:15]

set.seed(123)
num_row = length(Condition)
train_sample = sample(1:num_row, round(0.7*num_row), replace = FALSE)

colnames(tLIHC_data_sig) = gsub("-", ".", colnames(tLIHC_data_sig))
train_data = tLIHC_data_sig[train_sample,]

test_data = tLIHC_data_sig[-train_sample, ]

dim(train_data)
dim(test_data)
dim(tLIHC_data_sig)

LIHC_forest = randomForest(Condition ~ ., 
                           ntree = 100, 
                           data = train_data)
importance(LIHC_forest)[order(importance(LIHC_forest)),]
varImpPlot(LIHC_forest)

train_predictions = predict(LIHC_forest, train_data)
table(train_predictions, train_data$Condition)

test_predictions = predict(LIHC_forest, test_data)
table(test_predictions,test_data$Condition)
confusion_matrix = table(test_predictions, test_data$Condition)

# Calculate accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
cat("Model Accuracy: ", accuracy, "\n")
```


B. Change variables in the random forest run and check if the results are different. What are the parameters that you can achieve a better accuracy?


## Analysis

### miRNAs

What are your most significant miRNAs? Are they up-regulated or down-regulated? Are they highly correlated with each other?


### Genes

A. What are the common gene targets of your significant miRNAs?


B. Are any of your genes targeted by these top miRNAs known mutations of the disease you are researching?


## Conclusion

What do your findings suggest to you about potential diagnosis/treatment targets for the disease you are researching?



## Optional


A. Make a heatmap of all data (row clustering only). Use the colors darkblue, white and red2. [USA!USA!USA!].


B. find all correlations among the 30 most significant miRNAs for
1) all samples, 2) all tumor samples, 3) all control samples.
