---
title: Day1
author: Georgios Polyzotis
date: 
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## copyright (c)miRcore

## miRcore R camp day 1

(to add new chunk ctrl+alt+i or cmd+optn+i or green code button)

```{r}

```

### basic calculation

```{r}
a = 3
```

## vector

Take notes here!

```{r}
vec4 = 2:9
vec2 = c(2,3,4,5)

vec3 = seq(from=2, to =10, by=6)
```

### scan function

function scan() gets input from keyboards through Console window and
form a vector. Carriage return separates the element. Two carriage
return close the scan function. Default data type: character.

change input data type to numeric: scan(what=numeric()) 1) run the code:
vec4 \<- scan(what=numeric()) 2) go to Console section (lower left
section) and activate cursor there 3) type numbers 20 30 15 22 18 from
Console with carriage returns 4) end: one more carriage return

```{r}
scan(what=numeric())
vec4 = scan(what=numeric())

```

```{r}
vec4[1]
vec2[3] == 4
```

### element of vector []

```{r}
 #find the sum of the vector
sum(vec4)
 #find the mean of the vector
mean(vec4)
 #find the standard deviation of the vector
sd(vec4)

```

Check for logical values related to vec2:

```{r}
mix = c(1,2,3,4,"five")
mix = as.numeric(mix)

sum(mix, na.rm=T)
mean(mix,na.rm = T)
```

### vector has the same data type

```{r}

```

### special character

NA, NaN,

```{r}

```

## which

```{r}
which(vec4 == 20)

```

## matrix

What is a matrix? Definition here: two dimensional array

Method 1 of creating a matrix by using matrix() function

```{r}
M1 = matrix(2:7, nrow = 2, ncol = 3, byrow =T)
M1
# prints the matrix
```

Method 2 of creating a Matrix (by combining vectors)

```{r}
aVec = c(3,4,5,6)
bVec = c(3,5,7)

M2 = cbind(aVec,bVec)
M2

M2 = rbind(aVec,bVec)
M2[1,2]
M2
```

### matrix element

```{r}

m1 = matrix(c(10,21,3,14,9), nrow = 3, ncol = 2)
m1
vec1 = c(2,3,4,5,6)
which(m1 == 9, arr.ind = T)
length(vec1)
dim(m1)
str(m1)
```

## data frame

Let's make a data frame with: column1: numbers between 1 through 14
column2: letters a - d column 3: group 1001

```{r}
 #vector 1
numbers = 7:10
 #vector 2
letters = c("a","b","c","d")
 #vector 3
group = as.factor(c(1,0,0,1))
 #cbind three vectors
M = cbind.data.frame(numbers,letters,group)
 #what is the structure of mydata?
str(M)
 #let's add rownames
row.names(M) = c("poop", "pee", "fart", "cancer")

 #what does mydata look like now?
 #what are the column names of mydata
colnames(M)
 #what are the rownames of mydata
rownames(M) = c("pee","poop","cancer","fart")
```

Accessing rows/columns:

```{r}
M[2]
M[,2]
M["letters"]
M$letters

M[1,]
M["pee",]

```

Two ways to access a row

```{r}

```

Pattern match

```{r}
#shows top ~10 rows
head(M)
#shows bottom ~10 rows
tail(M)
#shows rows 1-3
M[1:3,]
#shows columns 1-3
M[,1:3]
#reorders col so you get group, letters, and numbers
M[,c("group","letters","numbers")]
# returns the row index of the letter column that has the value "b" (in this case, 2), so the full line prints the contents of that row
M[M[,"letters"]=="b",]
M[,"letters"] == "b"
M[3,2] = "c"
M[2,3]
M[1,]
vec = M[1:3,2]
vec
M[1:3,"numbers"]
M[M["numbers"]==9,2]
M[M["numbers"]==9,]

```
