---
title: "Graphs_Blank"
output: html_document
author: "Anna Kaehr"
date: 'July 2023'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today we will learn about how to make graphs to visually represent our data in a meaningful way

# miRcore R camp day 2-2
## graphs

random data

Lets make some random data to play with using the sample function. 
ex:We can choose 2 numbers out of 1,2,3,23 and 5 randomly without repeats.

```{r}

```

What if we want repeats? Then we need to use sampling with replacement.
This will choose 100 numbers randomly between 1 and 50, with repeats.

```{r}

```

What if we wanted to choose only even numbers?
ex: We can use sample() and seq() to choose 100 numbers out of the even numbers between 2 and 50 randomly with repeats.

```{r}

```


bargraph

Lets see if we can make a barplot of some data. The function we use is barplot(x) where x is a matrix or a vector. 

Our example below shows a barplot of 100 numbers chosen randomly between 1 and 50 with replacement.

```{r}

```
Lets add some arguments to add titles and axis labels.

```{r}

```


## add more argument
Lets try each argument to see the changes:

```{r}

```

Other arguments you can use:
legend.text=x, where x is a vector of strings* that are put in a legend
col=x, where x is a vector of colors for the bars and the legend
border=x, where x is a vector of colors for the borders of the bars
width=x, where x is a vector that changes the widths of the bars
space=x, where x is a ratio of each bar to be a space to the next bar
names.arg=x, where x is a vector of names for the bars
horiz=True/False, setting the bar graph to a horizontal or vertical one
density=25 changes the shading of the bars
angle=x, where x is a number of degrees for the shading (works with density argument)
ylim=c(ymin,ymax), where ymin and ymax are ranges for y-axis
Some arguments, like names.arg, need to have a number of elements equal to the number of bars. Others, like col and border, will repeat if there’re less elements than bars.


multiple graphs:
Sometimes we will want to arrange multiple graphs. This can be done using the par() function.
The example below shows how we can arrange 2 figures in 2 rows and 1 column.

```{r}

```

To return to one figure in one window we use par() again.

```{r}

```

Some really nice color selections can be found here: http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf

Breakout Room Task 1:
-Make 2 figures arranged in 1 rows and 2 column
-Each bar graph figure of four teams’ random scores between 1 and 30
-Two games of titles “Game 1” and “Game 2”
-Add ylim argument to make the same y-axis for both figures
-Change colors for each bar using the Rcolor.pdf
-par(mfrow=c(1,1))




```{r}

```

```{r}

```

Box Plots:(slide 16) In descriptive statistics, a boxplot is a convenient way of graphically depicting groups of numerical data through their quartiles. They may also have lines extending vertically from the boxes (whiskers) to indicate the variability outside the upper and lower quartiles. "Box and Whisker plot"
outliers may also be plotted as individual points 

Quartiles are defined by the following:
Minimum
Second quartile: the value below which the lower 25% of data is contained
Median: The middle number in a range of numbers
Third quartile: The value above which the upper 25% of data is contained
Maximum

##box plot

Lets make a boxplot using the boxplot() function! Use a sample of 1000 points betweeen 1 and 100 with replacement

```{r}

```
real data boxplot

Now lets look at a boxplot using real data- 
First make sure we have properly read in small_data.txt
```{r}

```

Lets make a boxplot for the Tumor and control samples 1 through 3.
what do you notice? How might we make a more visually informative plot?

```{r}

```
why log?
For high-throughput sequencing data, there is often a lot of data with very low expression values. Because the data is so skewed, it is hard to get an accurate view of the distribution of the data.

Log transformation can be used to make highly skewed data less skewed: 


```{r}

```
Why do we add a +1 to the data when making a log plot?

```{r}

```

```{r}

```

Grep reminder for task 1:
we need to make the following vectors for task 1: 
vaa should be all the samples for a miR_a
vta should be the Tumor samples for a miR_a
vca should be the control samples for a miR_a

Let's select row names with specific words such as "miR_a"
```{r}

```
```{r}

```



Task 1: 
-Make a matrix smalldataM from smalldata
-Make a bar graph of the miR_a values of all samples (vaa).
-Make a bar graph of the miR_a values of all tumor samples (vta).
-Make a bar graph of the miR_a values of all control samples (vca).
-Add title and x-axis and y-axis labels.
-Put graphs from 3 and 4 having the same y-axis range in one window.


answers:
1: Make a matrix smalldataM from smalldata
```{r}

```
make a bar graph of the miR_a values of all samples (vaa)
```{r}

```

3:make a bar graph of the miR_a values of all tumor samples
```{r}

```

4:Make a bar graph of all the control samples (vca)
```{r}

```

We can also do this using grep:
```{r}



```

Put graphs from 3 and 4 with the same y-axis range in one window
```{r}

```


## scatter plot
A scatter graph plots paired values. Here we'll use the function plot(x,y) where x and y are vectors.

```{r}


```
```{r}

```
We can set the x and y limits of the plots using: xlim() and ylim()
```{r}

```
You can find a variety of other graphic parameters such as line width and line type at this resource:
http://www.statmethods.net/advgraphs/parameters.html
lwd: line width
lty: line type


Task 2: Breakout Rooms!

Make a scatter plot of miR_h values between each patient’s tumor and control samples (tumor values on x-axis and control values on y-axis).


*If you don’t want scientific notation, give a penalty to it.
options(scipen=5)  # positive number prefers fixed value, negative number prefers scientific notation.


answer:

```{r}



```


We can use scatter plots to compare the expression values for different miRs. Lets look at miR_a plotted against miR_t
```{r}

```

```{r}

```
Quick break: 15 minutes



## correlation

See slides.

Correlation is a mutual relationship between two or more things. In statistics, you can think of correlation as a measure of how strongly two or more things are related.
Two variables with very high correllation will behave very similarly moving together.
Variables with correlation close to zero will not move together.


Two variables are said to be positively correlated if an increase in one is associated with an increase in the other
Two variables are said to be negatively correlated if an increase in one is associated with a decrease in the other

Shoe size and length of foot are very highly positively correlated: an increase in foot length will very strongly be associated with a larger shoe size

SAT score and GPA are somewhat positively correlated: People with higher GPA tend to score higher on the SAT, but not always

Shoe size and SAT score are uncorrelated (correlation close to zero): Think, do people with larger feet tend to do better/worse on the SAT?No, they tend to perform the same as people with small feet


Shoe size and length of foot are very highly positively correlated: an increase in foot length will very strongly be associated with a larger shoe size

SAT score and GPA are somewhat positively correlated: People with higher GPA tend to score higher on the SAT, but not always

Shoe size and SAT score are uncorrelated (correlation close to zero): Think, do people with larger feet tend to do better/worse on the SAT?No, they tend to perform the same as people with small feet

cor(x,y) calculates the correlation between the two vectors x and y
Values close to 1 or -1 imply strong correlation
Values close to 0 indicate little to no correlation


```{r}


```
Correlation does not imply causation!!

Example from wikipedia:Ice cream sales and the rate of drowning are positively correlated, but clearly ice cream consumption does not cause drowning.

The aforementioned example fails to recognize the importance of time and temperature in relationship to ice cream sales. Ice cream is sold during the hot summer months at a much greater rate than during colder times, and it is during these hot summer months that people are more likely to engage in activities involving water, such as swimming. The increased drowning deaths are simply caused by more exposure to water-based activities, not ice cream. 
These two variables are correlated, but do not cause each other




### Task using small_data.txt

Make vectors for each item

Make a scatter plot between miR_a and miR_c for all samples.
Make a scatter plot between miR_a and miR_k for all samples.
What is the correlated value between miR_a and miR_c?
What is the correlated value between miR_a and miR_k?

Answers:

```{r}


```

You downloaded your group data from Data folder
Read full_BRCA_miR_data.txt as fullData.
Get miR data distribution for each sample (hint: use boxplot function after log transform the data).
Get miR data distribution of first 15 samples.
Are the data comparable across samples?

Full Data:
```{r}

```


