---
title: "3_1 ttest"
author: "Prasanna Padmanabham"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## T-test
First, we need generate data that we can analyze
```{r}

```


Let's visualize the data we just generated

Now, let's run a ttest 
But before we do, how confident are that these distributions are similar (green check) or not similar (red cross)

```{r}

```

what if we just want the p.value, standard deviation or just the mean?
```{r}

```



Let's try this with small data! Run a t.test on the miR_a values for Tumor samples and Control samples
* if you forgot how use the functions for the mean (mean()) and standard deviation (sd()), feel free to use google or the help feature on R!

* R is a little quirky in that it can't find the mean of a row from a dataframe. To solve this, change it to a matrix (as.matrix()) or a vector (as.vector())

1. What are mean and standard deviation of miR_a values of Tumor samples?
```{r}
a = smalldata["miR_a",grep("Tumor",colnames(smalldata))]
a=as.numeric(a)
mean(a)
sd(a)
```

2. What are mean and standard deviation of miR_a values of Control samples?
```{r}
at = smalldata["miR_a",grep("Control",colnames(smalldata))]
at=as.numeric(at)
mean(at)
sd(at)
```

3. What is the p-value of student t-test between miR_a values of Tumor samples and miR_a values of Control samples?
```{r}
t.test(a,at)$p.value
```


4. Are the miR_a values of Tumor samples different from the miR_a values of Control samples? How confident are you?

Very confident that they are different

5. What is the log2 value of mean(miR_a of Tumor)/mean(miR_a of Control)?
```{r}
```

6. Is miR_a upregulated or downregulated in Tumor compared to Control? 


